---
title: "Multiple regression (childhood maltreatment & social, psychological & health traits)"
author: "Abigail ter Kuile"
date: "14/01/2022"
output: html_document
---

Resources: https://github.com/GenomicSEM/GenomicSEM/wiki/3.-Models-without-Individual-SNP-effects

#Clear workspace
```{r clear workspace}
rm(list = ls())
```

```{r load required packages}
require(GenomicSEM)
library(devtools)
require(Matrix)
require(stats)
```

#Run multivariable LDSC to create the S and V matrices 
```{r run multivariable LDSC, create S V matrices}
# load in munged summary statistics
traits <- c(
"data/CTS_noMHC.sumstats.gz", # Childhood maltreatment (PMID: 33740410)
"data/TIRE01_noMHC.sumstats.gz", # self-reported tiredness (PMID: 28322280)
"data/SUBJ01_noMHC.sumstats.gz", # subjective well-being (PMID: 27089181)
"data/RISK03_noMHC.sumstats.gz", # number of sexual partners (PMID: 30643258)
"data/RISK01_noMHC.sumstats.gz", # general risk tolerance (PMID: 30643258)
"data/INCO03_noMHC.sumstats.gz", # social deprivation (PMID: 27818178)
"data/HEAL01_noMHC.sumstats.gz", # self-reported health (PMID: 27864402)
"data/REPR01_noMHC.sumstats.gz", # age at first birth (PMID: 27798627)
"data/SMOK06_noMHC.sumstats.gz", # ever smoker (PMID: 30643258)
"data/NEUR02B_noMHC.sumstats.gz", # neuroticism (PMID: 30593698)
"data/INSO02_noMHC.sumstats.gz", # Insomnia (PMID: 30804565)
"data/LONG04_noMHC.sumstats.gz") # Parental attained age (PMID:27015805)

#Sample prevalence
sample.prev <- c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, .28, NA)

#Population prevalence
population.prev <- c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, .30, NA)

#the folder of LD scores
ld <- ("/data/eur_w_ld_chr/")
#the folder of LD weights
wld <- ("/data/eur_w_ld_chr/")

# Set trait names
trait.names<-c("CTS", "Tiredness", "Subj. wellbeing", "Sexual partners", "Risk", "Social dep", "Health", "Age first birth", "Ever smoker", "Neuroticism", "Insomnia", "Parents attained age")

#run LDSC
social_dem_traits_cts <- ldsc(traits, sample.prev, population.prev, ld, wld, trait.names, ldsc.log="social_dem_traits_cts")
``` 

optional command to save the output as a .RData file for later use
```{r save the ldsc output}
save(social_dem_traits_cts, file="social_dem_traits_cts.RData")
```

load the output if using later
```{r load ldsc output if needed}
load("social_dem_traits_cts.RData")
```

##1st TRAIT: add Tiredness. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness
```{r add tiredness specify model}
Tiredness_model <- 'CTS ~ Tiredness'
```

#run the model using the user defined function
```{r run +Tiredness model}
Tiredness_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = Tiredness_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +Tiredness}
Tiredness_output
```

##2nd TRAIT: add Subjective well-being. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness + Subjective well-being
```{r add Subjective well-being specify model}
subjwell_model <- 'CTS ~ Tiredness + Subj. wellbeing
Tiredness ~~ Subj. wellbeing'
```

#run the model using the user defined function
```{r run +subjwell model}
subjwell_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = subjwell_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +subjwell}
subjwell_output
```

##3rd TRAIT: add Social deprivation. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness + Subjective well-being + Social dep
```{r add Social dep specify model}
Socialdep_model <- 'CTS ~ Tiredness + Subj. wellbeing + Social dep
Tiredness ~~ Subj. wellbeing
Social dep ~~ Tiredness
Social dep ~~ Subj. wellbeing'
```

#run the model using the user defined function
```{r run +Socialdep model}
Socialdep_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = Socialdep_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +Socialdep}
Socialdep_output
```

##4th TRAIT: add number of sexual partners. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness + Subjective well-being + Social dep + number of sexual partners
```{r add sexualpartners specify model}
sexualpartners_model <- 'CTS ~ Tiredness + Subj. wellbeing + Social dep + Sexual partners
Tiredness ~~ Subj. wellbeing
Social dep ~~ Tiredness
Social dep ~~ Subj. wellbeing
Sexual partners ~~ Tiredness
Sexual partners ~~ Subj. wellbeing
Sexual partners ~~ Social dep
'
```

#run the model using the user defined function
```{r run +sexualpartners model}
sexualpartners_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = sexualpartners_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +sexualpartners}
sexualpartners_output
```

##5th TRAIT: add number of self-rated health. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness + Subjective well-being + Social dep + number of sexual partners + self rated health
```{r add Health specify model}
health_model <- 'CTS ~ Tiredness + Subj. wellbeing + Social dep + Sexual partners + Health
Tiredness ~~ Subj. wellbeing
Social dep ~~ Tiredness
Social dep ~~ Subj. wellbeing
Sexual partners ~~ Tiredness
Sexual partners ~~ Subj. wellbeing
Sexual partners ~~ Social dep
Health ~~ Tiredness
Health ~~ Subj. wellbeing
Health ~~ Social dep
Health ~~ Sexual partners
'
```

#run the model using the user defined function
```{r run +health model}
health_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = health_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +health}
health_output
```

##6th TRAIT: add general risk tolerance. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness + Subjective well-being + Social dep + number of sexual partners + self rated health + risk
```{r add Risk specify model}
Risk_model <- 'CTS ~ Tiredness + Subj. wellbeing + Social dep + Sexual partners + Health + Risk
Tiredness ~~ Subj. wellbeing
Social dep ~~ Tiredness
Social dep ~~ Subj. wellbeing
Sexual partners ~~ Tiredness
Sexual partners ~~ Subj. wellbeing
Sexual partners ~~ Social dep
Health ~~ Tiredness
Health ~~ Subj. wellbeing
Health ~~ Social dep
Health ~~ Sexual partners
Risk ~~ Tiredness
Risk ~~ Subj. wellbeing
Risk ~~ Social dep
Risk ~~ Sexual partners
Risk ~~ Health
'
```

#run the model using the user defined function
```{r run +Risk model}
Risk_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = Risk_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +Risk}
Risk_output
```


##7th TRAIT: add age first birth. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness + Subjective well-being + Social dep + number of sexual partners + self rated health + risk + age first birth
```{r 2 add birth specify model}
birth_model2 <- 'CTS ~ Tiredness + Subj. wellbeing + Social dep + Sexual partners + Health + Risk + Age first birth
Tiredness ~~ Subj. wellbeing
Social dep ~~ Tiredness
Social dep ~~ Subj. wellbeing
Sexual partners ~~ Tiredness
Sexual partners ~~ Subj. wellbeing
Sexual partners ~~ Social dep
Health ~~ Tiredness
Health ~~ Subj. wellbeing
Health ~~ Social dep
Health ~~ Sexual partners
Risk ~~ Tiredness
Risk ~~ Subj. wellbeing
Risk ~~ Social dep
Risk ~~ Sexual partners
Risk ~~ Health
Age first birth ~~ Tiredness
Age first birth ~~ Subj. wellbeing
Age first birth ~~ Social dep
Age first birth ~~ Sexual partners
Age first birth ~~ Health
Age first birth ~~ Risk
'
```

#run the model using the user defined function
```{r run +birth model}
birth_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = birth_model2, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +birth}
birth_output
```

##8th TRAIT: add neuroticism. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness + Subjective well-being + Social dep + number of sexual partners + self rated health + risk + age first birth + neurotisicsm
```{r 2 add Neuroticism specify model}
neuroticism_model2 <- 'CTS ~ Tiredness + Subj. wellbeing + Social dep + Sexual partners + Health + Risk + Age first birth + Neuroticism
Tiredness ~~ Subj. wellbeing
Social dep ~~ Tiredness
Social dep ~~ Subj. wellbeing
Sexual partners ~~ Tiredness
Sexual partners ~~ Subj. wellbeing
Sexual partners ~~ Social dep
Health ~~ Tiredness
Health ~~ Subj. wellbeing
Health ~~ Social dep
Health ~~ Sexual partners
Risk ~~ Tiredness
Risk ~~ Subj. wellbeing
Risk ~~ Social dep
Risk ~~ Sexual partners
Risk ~~ Health
Age first birth ~~ Tiredness
Age first birth ~~ Subj. wellbeing
Age first birth ~~ Social dep
Age first birth ~~ Sexual partners
Age first birth ~~ Health
Age first birth ~~ Risk
Neuroticism ~~ Tiredness
Neuroticism ~~ Subj. wellbeing
Neuroticism ~~ Social dep
Neuroticism ~~ Sexual partners
Neuroticism ~~ Health
Neuroticism ~~ Risk
Neuroticism ~~ Age first birth
'
```

#run the model using the user defined function
```{r run +neuroticism model}
neuroticism_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = neuroticism_model2, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +neuroticism}
neuroticism_output
``` 

##9th TRAIT: add insomnia. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness + Subjective well-being + Social dep + number of sexual partners + self rated health + risk + age first birth + neurotisicsm + insomnia
```{r 2 add insomnia specify model}
insomnia_model2 <- 'CTS ~ Tiredness + Subj. wellbeing + Social dep + Sexual partners + Health + Risk + Age first birth + Neuroticism + Insomnia
Tiredness ~~ Subj. wellbeing
Social dep ~~ Tiredness
Social dep ~~ Subj. wellbeing
Sexual partners ~~ Tiredness
Sexual partners ~~ Subj. wellbeing
Sexual partners ~~ Social dep
Health ~~ Tiredness
Health ~~ Subj. wellbeing
Health ~~ Social dep
Health ~~ Sexual partners
Risk ~~ Tiredness
Risk ~~ Subj. wellbeing
Risk ~~ Social dep
Risk ~~ Sexual partners
Risk ~~ Health
Age first birth ~~ Tiredness
Age first birth ~~ Subj. wellbeing
Age first birth ~~ Social dep
Age first birth ~~ Sexual partners
Age first birth ~~ Health
Age first birth ~~ Risk
Neuroticism ~~ Tiredness
Neuroticism ~~ Subj. wellbeing
Neuroticism ~~ Social dep
Neuroticism ~~ Sexual partners
Neuroticism ~~ Health
Neuroticism ~~ Risk
Neuroticism ~~ Age first birth
Insomnia ~~ Tiredness
Insomnia ~~ Subj. wellbeing
Insomnia ~~ Social dep
Insomnia ~~ Sexual partners
Insomnia ~~ Health
Insomnia ~~ Risk
Insomnia ~~ Age first birth
Insomnia ~~ Neuroticism
'
```

#run the model using the user defined function
```{r run +insomnia model}
insomnia_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = insomnia_model2, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +insomnia}
insomnia_output
``` 

##10th TRAIT: add smoking. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness + Subjective well-being + Social dep + number of sexual partners + self rated health + risk + age first birth + neurotisicsm + insomnia + Ever smoker
```{r 2 add smoker specify model}
smoker_model2 <- 'CTS ~ Tiredness + Subj. wellbeing + Social dep + Sexual partners + Health + Risk + Age first birth + Neuroticism + Insomnia + Ever smoker
Tiredness ~~ Subj. wellbeing
Social dep ~~ Tiredness
Social dep ~~ Subj. wellbeing
Sexual partners ~~ Tiredness
Sexual partners ~~ Subj. wellbeing
Sexual partners ~~ Social dep
Health ~~ Tiredness
Health ~~ Subj. wellbeing
Health ~~ Social dep
Health ~~ Sexual partners
Risk ~~ Tiredness
Risk ~~ Subj. wellbeing
Risk ~~ Social dep
Risk ~~ Sexual partners
Risk ~~ Health
Age first birth ~~ Tiredness
Age first birth ~~ Subj. wellbeing
Age first birth ~~ Social dep
Age first birth ~~ Sexual partners
Age first birth ~~ Health
Age first birth ~~ Risk
Neuroticism ~~ Tiredness
Neuroticism ~~ Subj. wellbeing
Neuroticism ~~ Social dep
Neuroticism ~~ Sexual partners
Neuroticism ~~ Health
Neuroticism ~~ Risk
Neuroticism ~~ Age first birth
Insomnia ~~ Tiredness
Insomnia ~~ Subj. wellbeing
Insomnia ~~ Social dep
Insomnia ~~ Sexual partners
Insomnia ~~ Health
Insomnia ~~ Risk
Insomnia ~~ Age first birth
Insomnia ~~ Neuroticism
Ever smoker ~~ Tiredness
Ever smoker ~~ Subj. wellbeing
Ever smoker ~~ Social dep
Ever smoker ~~ Sexual partners
Ever smoker ~~ Health
Ever smoker ~~ Risk
Ever smoker ~~ Age first birth
Ever smoker ~~ Neuroticism
Ever smoker ~~ Insomnia
'
```

#run the model using the user defined function
```{r run +smoker model}
smoker_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = smoker_model2, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +smoker}
smoker_output
``` 

##11th TRAIT: add parents attained age. Specify a model in which CTS is regressed on top rg trait: self-reported tiredness + Subjective well-being + Social dep + number of sexual partners + self rated health + risk + household income + age first birth + neurotisicsm + insomnia + Ever smoker + parents attained age 
```{r 2 add parentsage specify model}
parentsage_model2 <- 'CTS ~ Tiredness + Subj. wellbeing + Social dep + Sexual partners + Health + Risk + Age first birth + Neuroticism + Insomnia + Ever smoker + Parents attained age
Tiredness ~~ Subj. wellbeing
Social dep ~~ Tiredness
Social dep ~~ Subj. wellbeing
Sexual partners ~~ Tiredness
Sexual partners ~~ Subj. wellbeing
Sexual partners ~~ Social dep
Health ~~ Tiredness
Health ~~ Subj. wellbeing
Health ~~ Social dep
Health ~~ Sexual partners
Risk ~~ Tiredness
Risk ~~ Subj. wellbeing
Risk ~~ Social dep
Risk ~~ Sexual partners
Risk ~~ Health
Age first birth ~~ Tiredness
Age first birth ~~ Subj. wellbeing
Age first birth ~~ Social dep
Age first birth ~~ Sexual partners
Age first birth ~~ Health
Age first birth ~~ Risk
Neuroticism ~~ Tiredness
Neuroticism ~~ Subj. wellbeing
Neuroticism ~~ Social dep
Neuroticism ~~ Sexual partners
Neuroticism ~~ Health
Neuroticism ~~ Risk
Neuroticism ~~ Age first birth
Insomnia ~~ Tiredness
Insomnia ~~ Subj. wellbeing
Insomnia ~~ Social dep
Insomnia ~~ Sexual partners
Insomnia ~~ Health
Insomnia ~~ Risk
Insomnia ~~ Age first birth
Insomnia ~~ Neuroticism
Ever smoker ~~ Tiredness
Ever smoker ~~ Subj. wellbeing
Ever smoker ~~ Social dep
Ever smoker ~~ Sexual partners
Ever smoker ~~ Health
Ever smoker ~~ Risk
Ever smoker ~~ Age first birth
Ever smoker ~~ Neuroticism
Ever smoker ~~ Insomnia
Parents attained age ~~ Tiredness
Parents attained age ~~ Subj. wellbeing
Parents attained age ~~ Social dep
Parents attained age ~~ Sexual partners
Parents attained age ~~ Health
Parents attained age ~~ Risk
Parents attained age ~~ Age first birth
Parents attained age ~~ Neuroticism
Parents attained age ~~ Insomnia
Parents attained age ~~ Ever smoker
'
```

#run the model using the user defined function
```{r run +parentsage model}
parentsage_output <- usermodel(covstruc = social_dem_traits_cts, estimation = "DWLS", model = parentsage_model2, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +parentsage}
parentsage_output
``` 

