---
title: "Multiple regression (childhood maltreatment & psychiatric disorders)"
author: "Abigail ter Kuile"
date: "14/01/2022"
output: html_document
---

Resources: https://github.com/GenomicSEM/GenomicSEM/wiki/3.-Models-without-Individual-SNP-effects

#Clear workspace
```{r clear workspace}
rm(list = ls())
```

```{r load required packages}
require(GenomicSEM)
library(devtools)
require(Matrix)
require(stats)
```


# 2.Run multivariable LDSC to create the S and V matrices
```{r run multivariable LDSC, create S V matrices}
# load in munged summary statistics
traits <- c("data/CTS_noMHC.sumstats.gz", # Childhood maltreatment (PMID: 30478444)
            "data/ADHD05_noMHC.sumstats.gz", # ADHD (PMID: 30478444)
            "data/ALCD03_noMHC.sumstats.gz", # Alcohol dependence (PMID: 30482948)
            "data/ANXI03_noMHC.sumstats.gz", # Any lifetime anxiety disorder (PMID: 31748690)
            "data/AUTI07_noMHC.sumstats.gz", # Autism spectrum disorder (PMID: 30804558)
            "data/DEPR01_noMHC.sumstats.gz", # Major depressive disorder (PMID: 29700475)
            "data/SCHI02_noMHC.sumstats.gz", # Schizophrenia (PMID: 25056061)
            "data/TotalPCL_MVP_eur_noMHC.sumstats.gz") # PTSD symptoms (PMID: 33510476)

#Sample prevalence
sample.prev <- c(NA,.36,.25,.30,.40,.28,.44,NA)

#Population prevalence
population.prev <- c(NA,.05,.159,.16,.01,.15,.01,NA)

#the folder of LD scores
ld <- ("data/eur_w_ld_chr/")
#the folder of LD weights
wld <- ("data/eur_w_ld_chr/")

# Set trait names
trait.names<-c("CTS",
               "ADHD",
               "ALCH",
               "ANX",
               "ASD",
               "MDD",
               "SCZ",
               "PTSD")

#run LDSC
PD_regression_cts_ptsd_tot <- ldsc(traits,
                                   sample.prev,
                                   population.prev,
                                   ld,
                                   wld,
                                   trait.names,
                                   ldsc.log="PD_regression_cts_ptsd_tot")
```

optional command to save the output as a .RData file for later use
```{r save the ldsc output}
save(PD_regression_cts_ptsd_tot, file="PD_regression_cts_ptsd_tot.RData")
```

load the output if using later
```{r load the ldsc output}
load(file="PD_regression_cts_ptsd_tot.RData")
```

###TRAIT 1 ADHD: specify a model in which CTS is regressed on the top rG trait (ADHD)
```{r trait 1 specify PTSD PTSD model}
adhd_model <- 'CTS ~ ADHD'
```

#run the model using the user defined function
```{r run PTSD model}
adhd_output <- usermodel(covstruc = PD_regression_cts_ptsd_tot, estimation = "DWLS", model = adhd_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output PTSD}
adhd_output
```

###TRAIT 2: add PTSD. Specify a model in which CTS is regressed on ADHD + PTSD
```{r specify model add adhd}
PTSD_model <- 'CTS ~ ADHD + PTSD
ADHD ~~ PTSD'
```

#run the model using the user defined function
```{r run +adhd model}
PTSD_output <- usermodel(covstruc = PD_regression_cts_ptsd_tot, estimation = "DWLS", model = PTSD_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output PTSD}
PTSD_output
```

###3RD TRAIT: add MDD. Specify a model in which CTS is regressed on ADHD + PTSD + MDD
```{r add MDD specify model}
MDD_model <- 'CTS ~ ADHD + PTSD + MDD
PTSD ~~ ADHD
PTSD ~~ MDD
MDD ~~ ADHD'
```

#run the model using the user defined function
```{r run +MDD model}
MDD_output <- usermodel(covstruc = PD_regression_cts_ptsd_tot, estimation = "DWLS", model = MDD_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output
```{r output +mdd}
MDD_output
```

###4TH TRAIT: add alcohol dependence. Specify a model in which CTS is regressed on ADHD + PTSD + MDD + ALCH
```{r +ALCH specify model}
ALCH_model <- 'CTS ~ ADHD + PTSD + MDD + ALCH
PTSD ~~ ADHD
PTSD ~~ MDD
MDD ~~ ADHD
ALCH ~~ PTSD
ALCH ~~ ADHD
ALCH ~~ MDD
'
```

#run the model using the user defined function
```{r run +ALCH model}
ALCH_output <- usermodel(covstruc = PD_regression_cts_ptsd_tot, estimation = "DWLS", model = ALCH_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output 
```{r output +ALCH}
ALCH_output
```

###5TH TRAIT: add Anxiety disorder . Specify a model in which CTS is regressed on ADHD + PTSD + MDD + ALCH + ANX
```{r +ANX specify model}
ANX_model <- 'CTS ~ ADHD + PTSD + MDD + ALCH + ANX
PTSD ~~ ADHD
PTSD ~~ MDD
MDD ~~ ADHD
ALCH ~~ PTSD
ALCH ~~ ADHD
ALCH ~~ MDD
ANX ~~ PTSD
ANX ~~ ADHD
ANX ~~ MDD
ANX ~~ ALCH
'
```

#run the model using the user defined function
```{r run +ANX model}
ANX_output <- usermodel(covstruc = PD_regression_cts_ptsd_tot, estimation = "DWLS", model = ANX_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output 
```{r output +ANX}
ANX_output
```

###6TH TRAIT: add Autism spectrum disorder. Specify a model in which CTS is regressed on ADHD + PTSD + MDD + ALCH + ANX + ASD
```{r +ASD specify model}
ASD_model <- 'CTS ~ ADHD + PTSD + MDD + ALCH + ANX + ASD
PTSD ~~ ADHD
PTSD ~~ MDD
MDD ~~ ADHD
ALCH ~~ PTSD
ALCH ~~ ADHD
ALCH ~~ MDD
ANX ~~ PTSD
ANX ~~ ADHD
ANX ~~ MDD
ANX ~~ ALCH
ASD ~~ PTSD
ASD ~~ ADHD
ASD ~~ MDD
ASD ~~ ALCH
ASD ~~ ANX
'
```

#run the model using the user defined function
```{r run +ASD model}
ASD_output <- usermodel(covstruc = PD_regression_cts_ptsd_tot, estimation = "DWLS", model = ASD_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output 
```{r output +ASD}
ASD_output
```

###8TH TRAIT: add Schizophrenia. Specify a model in which CTS is regressed on ADHD + PTSD + MDD + ALCH + ANX + ASD + SCZ
```{r +SCZspecify model}
SCZ_model <- 'CTS ~ ADHD + PTSD + MDD + ALCH + ANX + ASD + SCZ
PTSD ~~ ADHD
PTSD ~~ MDD
MDD ~~ ADHD
ALCH ~~ PTSD
ALCH ~~ ADHD
ALCH ~~ MDD
ANX ~~ PTSD
ANX ~~ ADHD
ANX ~~ MDD
ANX ~~ ALCH
ASD ~~ PTSD
ASD ~~ ADHD
ASD ~~ MDD
ASD ~~ ALCH
ASD ~~ ANX
SCZ ~~ PTSD
SCZ ~~ ADHD
SCZ ~~ MDD
SCZ ~~ ALCH
SCZ ~~ ANX
SCZ ~~ ASD
'
```

#run the model using the user defined function
```{r run +SCZ model}
SCZ_output <- usermodel(covstruc = PD_regression_cts_ptsd_tot, estimation = "DWLS", model = SCZ_model, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)
```

#print the output 
```{r output +SCZ}
SCZ_output
```

